DeepSphere: a graph-based spherical CNN

### 2019-11-15 main.tex
better kernel width figure
first big pass on section 3 and around

\newcommand{\fix}{\marginpar{FIX}}
\newcommand{\new}{\marginpar{NEW}}
\newcommand{\todo}[1]{{\color[rgb]{.6,.1,.6}{#1}}}
\newcommand{\nati}[1]{{\color[rgb]{.3,.5,.9}{#1}}}
\newcommand{\mart}[1]{{\color[rgb]{.9,.5,.3}{#1}}}
\newcommand{\mdeff}[1]{{\color[rgb]{.3,.8,.3}{#1}}}

\newcommand{\figref}[1]{figure~\ref{fig:#1}}
\newcommand{\Figref}[1]{Figure~\ref{fig:#1}}
\newcommand{\tabref}[1]{table~\ref{tab:#1}}
\newcommand{\Tabref}[1]{Table~\ref{tab:#1}}
\newcommand{\secref}[1]{section~\ref{sec:#1}}
\newcommand{\Secref}[1]{Section~\ref{sec:#1}}
%\newcommand{\secref}[1]{\S\ref{sec:#1}}
\newcommand{\eqnref}[1]{equation~(\ref{eqn:#1})}
\newcommand{\Eqnref}[1]{Equation~(\ref{eqn:#1})}

@@

\todo{
TODO:
\begin{itemize}
	\item PSD of climate data: how does it compare to cosmo and 3D objects? Climate data on spectrum plot. \nati{I would drop this}
	\item Unify vocabulary. Sampling, discretization, or pixelization of the sphere? Signal, maps, fields, functions. Irregular, uneven. Coarsening, down-sampling. Sample, vertex, node, pixel, point.
	\item check English vs American English
	\item Explain heuristic used to set the kernel width in v1. Add values to \figref{kernel_widths} for comparison. \nati{Not sure what you mean here.}
	\item put all the code (learning experiments from Frédérick and equivariance experiments from Martino) in a common reproducible git repository to be linked from the paper
	\item update pygsp and merge sphere branch
	\item Compute the invariance error of the whole NNs (before and after training, invariance can be learned). Add it with equivariance error of the convolution to \tabref{cosmo}. \nati{I guess we are not doing that}
	\item insert \citet{jiang2019sphericalcnn} in intro: Need a global coordinate system (ok for planets, not projections like cosmo)
	\item mention that \citep{esteves2018sphericalcnn} and \citep{cohen2018sphericalcnn} computed equivariance error
\end{itemize}
}

\todo{
rebuttal:
\begin{itemize}
    \item answer commenter
	\item add somewhere (intro?): exploit symmetries to reduce sample cost $\Rightarrow$ we don't discuss symmetries much
	\item \citet{mudigonda2017climateevents} use Integrated Water Vapor (IWV) only. That most probably is the same as our TMQ, which is most probably the most informative feature. TODO: run ours with TMQ alone.
	\item Show the three samplings in a figure?
	\item better abstract (Friday's talk abstract?)
	\item conclusion: less speculation?
	\item check feedback from ICLR'20 LTS2 reading group
	\item \mdeff{check my Telegram notes}
\end{itemize}
}

@@

\todo{
Not done:
\begin{itemize}
	\item Do the experiments on the icosahedral grid with the pooling as defined in the paper (i.e., not Jiang's specialty of taking the value of a pre-chosen sub-pixel). \nati{I guess we are not doing that.} \mdeff{I'm not sure what is the state of this, and Frédérick is not with us anymore. It's quite minor.}
	\item Check results for ModelNet40 and integrate in text. \nati{I would drop this.} \mdeff{Yeah, those results are not mature enough.}
	\item Add \citet{bruna2013gnn} in equivariance? They were the first to do a spherical CNN with graphs. Graph is built from a uniform random sampling of the sphere. \nati{Where do you want to add them?} \mdeff{They don't write how their Laplacian was built. Also, they only did MNIST on the sphere...}
\end{itemize}
}

@@

\begin{abstract}
Designing a convolution for a spherical neural network requires a delicate tradeoff between efficiency and rotation equivariance.
% exactitude with respect to equivariance.
DeepSphere, a method based on a graph representation of the discretized sphere, strikes a controllable balance between these two desiderata.
% Say something here about isotropy?
This contribution is twofold.
First, we study both theoretically and empirically how equivariance is affected by the underlying graph with respect to the number of pixels and neighbors.
Second, we evaluate DeepSphere on relevant problems.
Experiments show state-of-the-art performance and demonstrates the efficiency and flexibility of this formulation.
Perhaps surprisingly, comparison with previous work suggests that anisotropic filters might be an unnecessary price to pay.
%Say something about speed here... Not sure what?
% memory, computations, implementation complexity
\end{abstract}

@@

\begin{figure}[h]
	\centering
	\subfloat[]{
		\includegraphics[height=0.20\linewidth,valign=t]{example_brain_meg}
	} \hspace{0.1em}
	\subfloat[]{
		\includegraphics[height=0.20\linewidth,valign=t]{example_cosmo_cmb}
	} \hfill
	\subfloat[]{
		\includegraphics[height=0.20\linewidth,valign=t]{example_climate_TMQ}
		\label{fig:examples:climate}
	} \hfill
	\subfloat[]{
		\includegraphics[height=0.20\linewidth,valign=t]{example_ghcn_daily_tmax}
		\label{fig:examples:ghcn:tmax}
	} \hfill
	\subfloat[]{
		\includegraphics[height=0.20\linewidth,valign=t]{example_ghcn_graph}
		\label{fig:examples:ghcn:graph}
	}
	\caption{
		Examples of spherical data:
		(a) brain activity recorded through magnetoencephalography (MEG),\protect\footnotemark
		(b) the cosmic microwave background (CMB) temperature from \citet{planck2015overview},
		(c) hourly precipitation from a climate simulation \citep{jiang2019sphericalcnn}, %\citep{mudigonda2017climateevents},
		(d) daily maximum temperature from the Global Historical Climatology Network (GHCN).\protect\footnotemark
		A rigid full-sphere pixelization is not ideal: brain activity is only measured on the scalp, the Milky Way's galactic plane masks observations, climate scientists desire a variable resolution, and the position of weather stations is arbitrary and changes over time.
		(e) Graphs can faithfully and efficiently represent sampled spherical data by placing vertices where it matters.
		%\todo{horizontal colorbar for brain}
	}
	\label{fig:examples}
\end{figure}
\footnotetext[1]{\scriptsize\url{https://martinos.org/mne/stable/auto_tutorials/plot_visualize_evoked.html}}
\footnotetext[2]{\scriptsize\url{https://www.ncdc.noaa.gov/ghcn-daily-description}}

@@

% keep the story simple and focus on the tradeoff between cost (time / computational and storage / memory) and exactitude
As neural networks (NNs) have proved to be great tools for inference, variants have been developed to handle spherical data.
Exploiting the locally Euclidean property of the sphere, early attempts used standard 2D convolutions on a grid discretization of the sphere \citep{boomsma2017sphericalcnn, su2017sphericalcnn, coors2018sphericalcnn}.
While simple and efficient, those convolutions are not equivariant to rotations.
On the other side of this tradeoff, \citet{cohen2018sphericalcnn} and \citet{esteves2018sphericalcnn} proposed to perform proper spherical convolutions through the spherical harmonic transform.
While equivariant to rotations, those convolutions are expensive (\secref{method:convolution}).

@@

As a lack of equivariance can penalize performance (\secref{exp:cosmo}) and expensive convolutions prohibit their application to some real-world problems, methods standing between these two extremes are desired.
%Such high resolutions are not necessary to represent the depth maps of 3D objects, but are required to faithfully represent cosmological and climate data.
\citet{cohen2019gauge} proposed to reduce costs by limiting the size of the representation of the symmetry group by projecting the data from the sphere to the icosahedron.
The distortions introduced by this projection might however hinder performance (\secref{exp:climate}).
% (from the continuous rotation group SO(3) to a discrete rotation group)
% the dimensionality of feature maps is equal to the size of the irreducible representation of the symmetry group
%As a platonic solid, it \todo{exhibits discrete rotations}.
% the representation of its symmetry group is much smaller than SO(3)
%\todo{not all rotations, subset of SO(3)?}

Another approach is to represent the discretized sphere as a graph connecting pixels according to the distance between them \citep{bruna2013gnn, khasanova2017sphericalcnn, perraudin2019deepspherecosmo}.
While Laplacian-based graph convolutions are more efficient than spherical convolutions, they are not exactly equivariant \citep{defferrard2019deepsphereequiv}.
In this work, we argue that graph-based spherical CNNs strike an interesting balance, with a controllable tradeoff between cost and equivariance (which is linked to performance).
Experiments on multiple problems of practical interest show the competitiveness and flexibility of this approach.
% Furthermore, experiments show that, perhaps surprisingly and contrarily to previously published results \citep{cohen2019gauge} \todo{[Bronstein?]}, anisotropic filters might be an unnecessary price to pay.
% This paper focuses on improving the rotation equivariance of \citet{perraudin2019deepspherecosmo}, a scalable graph-based method designed for cosmological applications.

@@

DeepSphere leverages graph convolutions to achieve the following properties: (i) computational efficiency, (ii) sampling flexibility, and (iii) rotation equivariance (\secref{equivariance}).
% An example architecture is shown in \figref{architecture}.
The main idea is to model the discretized sphere as a graph of connected pixels: the length of the shortest path between two pixels is an approximation of the geodesic distance between them.
We use the graph CNN formulation introduced in \citep{defferrard2016graphnn} and a pooling strategy that exploits hierarchical pixelizations of the sphere.

\paragraph{Sampling.}
Given a sampling scheme $\V = \{x_i \in \S^2\}_{i=1}^n$, a function $f : \S^2 \supset F_\V \to \R$ is sampled as $T_\V(f) = \b{f}$ by the sampling operator $T_\V: L^2(\S^2) \supset F_\V \to \R^n$ defined as $\b{f}: f_i=f(x_i)$, where $F_\V$ will be specified in \secref{equivariance}.
As there is no analogue of uniform sampling on the sphere, many schemes have been proposed with different tradeoffs.
In this work, depending on the considered application, we will use the equiangular \citep{driscoll1994Fouriersphere}, HEALPix \citep{gorski2005healpix}, and icosahedral \citep{baumgardner1985icosahedral} grids.

\paragraph{Graph.}
From $\V$, we construct a weighted undirected graph $\G = (\V, w)$, where the elements of $\V$ are the vertices and the weight $w_{ij} = w_{ji}$ is a similarity measure between vertices $x_i$ and $x_j$.
The combinatorial graph Laplacian $\b{L} \in \R^{n \times n}$ is defined as $\b{L} = \b{D} - \b{A}$, where $\b{A} = (w_{ij})$ is the weighted adjacency matrix, $\b{D} = (d_{ii})$ is the diagonal degree matrix, and $d_{ii} = \sum_j w_{ij}$ is the weighted degree of vertex $x_i$.
%$\b{D} = \diag(\b{A} \b{1})$
Given a sampling scheme $\V$, usually fixed by the application or the available measurements, the freedom in constructing $\G$ is in setting $w$.
\Secref{equivariance} shows how to set $w$ to minimize the equivariance error.
%building a good graph amounts to set a good $w$.

\paragraph{Convolution.} \label{sec:method:convolution}
On Euclidean domains, convolutions are efficiently implemented by sliding a window in the signal domain.
On the sphere however, there is no straightforward way to implement a convolution in the signal domain due to non-uniform samplings.
Convolutions are most often performed in the spectral domain through a spherical harmonic transform (SHT).
That is the approach taken by \citet{cohen2018sphericalcnn} and \citet{esteves2018sphericalcnn}, which has a computational cost of $\bO(n^{3/2})$ on isolatitude pixelizations (such as the HEALPix and equiangular grids) and $\bO(n^2)$ in general.
On the other hand, following \citet{defferrard2016graphnn}, graph convolutions can be defined as
\begin{equation} \label{eqn:graph_conv}
	h(\b{L}) \b{f} = \left(\sum_{i=0}^K \alpha_i \b{L}^i\right) \b{f},
\end{equation}
where $K$ is the polynomial order (which corresponds to the filter's size) and $\alpha_i$ are the coefficients to be optimized during training.\footnote{In practice, training with Chebyshev polynomials (instead of monomials) is slightly more stable. We believe it to be due to their orthogonality and uniformity.}
Those convolutions are used by \citet{khasanova2017sphericalcnn} and \citet{perraudin2019deepspherecosmo} and cost $\bO(n)$ operations through a recursive application of $\b{L}$.\footnote{As long as the graph is sparsified such that the number of edges, i.e., the number of non-zeros in $\b{A}$, is proportional to the number of vertices $n$. This can always be done as most weights are very small.}

\paragraph{Pooling.}
Down- and up-sampling is natural for hierarchical pixelizations,\footnote{The equiangular, HEALPix, and icosahedral schemes are of this kind.} where each subdivision divides a pixel in (an equal number of) child sub-pixels.
To pool (down-sample), the data supported on the sub-pixels is summarized by a permutation invariant function such as the maximum or the average.
To unpool (up-sample), the data supported on a pixel is copied to all its sub-pixels.
% Frédérick: Dans le cas de l'icosahedron, le pooling n'est pas invariant à la permutation (on prend la valeur du 1er pixel, sans regarder les autres)
% Frédérick: Et le unpooling est correct uniquement dans le cas de l'average pooling (mais on fait pas de max pooling dans ce cas). Encore une fois, l'icosahedron est inspiré de Jiang, où 1 seul sub-pixel prend la valeur du pixel, le reste est rempli avec des '1'

@@

\section{Graph convolution and equivariance} \label{sec:equivariance}
%\section{Is graph convolution equivariant to rotation? [2 pages]} \label{sec:equivariance}

@@

We require $F_\V$ to be a suitable subspace of $L^2(\S^2)$ such that $T_\V$ is invertible, i.e., the function $f \in F_\V$ can be unambiguously reconstructed from its sampled values $\b{f}$.
The existence of such a subspace depends on the sampling scheme $\V$, and its characterization is a common problem in signal processing \citep{driscoll1994Fouriersphere}.
For most sampling schemes, it is not known if $F_\V$ exists and hence if $T_\V$ is invertible.
A special case is the equiangular sampling scheme where a sampling theorem holds, and thus a closed form of $T_\V^{-1}$ is known. % \citep{driscoll1994Fouriersphere}
For sampling schemes where no such sampling formula is available, we leverage the discrete SHT to reconstruct $f$ from $\b{f}=T_\V f$, thus approximating $T_\V^{-1}$.
For all theoretical considerations, we assume that $F_\V$ exists and $f \in F_\V$.

% 0.2 Explain why the Laplacian is so important...
By definition, the (spherical) graph convolution is rotation equivariant, if and only if, it commutes with the rotation operator defined as $R(g), g\in SO(3)$: $R(g) f(x) = f\left(g^{-1} x \right)$.
In the context of this work, graph convolution is performed by recursive applications of the graph Laplacian (\eqnref{graph_conv}).
Hence, if $\b{L}$ commutes with the rotation operator, then, by recursion, it will also commute with the convolution $h(\b{L})$.
As a result, to simplify our analysis, we concentrate on the graph Laplacian and $h(\b{L})$ is rotational invariant if and only if
% 0.3 Formulate the problem
\begin{equation} \label{eq:equivariance}
	\b{R}_\V(g) \b{L} \b{f} = \b{L} \b{R}_\V(g) \b{f} \hspace{1cm} \forall f\in F_\V \text{ and } \forall g\in SO(3),
\end{equation}
where $\b{R}_\V(g) = T_\V R(g) T_\V^{-1}$. For an empirical evaluation of the equivariance, we define the \textit{normalized equivariance error} for a signal $\b{f}$, and a  rotation $g$ as:
\begin{equation} \label{eq:equivariance error}
	E_{\b{L}}(\b{f}, g) = \left(\frac{ \norm {\b{R}_\V(g) \b{L} \b{f} - \b{L} \b{R}_\V(g) \b{f}} }{\norm {\b{L} \b{f}}}\right)^2,
\end{equation}
More generally for a class of signal $f \in C \subset F_\V$, the \textit{mean equivariance error} defined as
\begin{equation} \label{eq:mean equivariance error}
	\overline E_{\b{L}, C} = \mathbb E_{\b{f}\in C, g\in SO(3)} \ E_{\b{L}}(\b{f}, g),
\end{equation}
represents well the overall equivariance error.
In practice, we analyse this error with respect of the frequency by setting the set $C$ to functions $f$ made of single degree spherical harmonic.
The expected value is obtained by averaging over a finite number of random functions and random rotations.

\subsection{Finding the optimal set of weights}  \label{sec:optimal}
% 1) summary the work of Khasanova
Naturally, \citet{khasanova2017sphericalcnn} designed a Laplacian $\b{L}$ that minimizes \eqref{eq:equivariance error} for specific rotations on the longitude and latitude axis.
Working specifically with the equiangular sampling, they consider the set $\mathcal{G}$ of all the possible graphs where each node is connected only to four of its nearest neighbours (North, South, West, East).
% and propose a weighting scheme $w_{ij}$ that minimizes \eqref{eq:equivariance error} for specific rotations on the longitude latitude axis.
Their approach leads to weights $w_{ij}$ inversely proportional to the Euclidean distance between vertices:
\begin{equation} \label{eqn:weights:khasanova}
	w_{ij} = \frac{1}{\norm{x_i-x_j}}
\end{equation}

>>>>

\newcommand{\fix}{\marginpar{FIX}}
\newcommand{\new}{\marginpar{NEW}}
\newcommand{\todo}[1]{{\color[rgb]{.6,.1,.6}{#1}}}
\newcommand{\nati}[1]{{\color[rgb]{.3,.5,.9}{#1}}}
\newcommand{\mart}[1]{{\color[rgb]{.9,.5,.3}{#1}}}
\newcommand{\mdeff}[1]{{\color[rgb]{.8,.3,.2}{#1}}}

\newcommand{\figref}[1]{figure~\ref{fig:#1}}
\newcommand{\Figref}[1]{Figure~\ref{fig:#1}}
\newcommand{\tabref}[1]{table~\ref{tab:#1}}
\newcommand{\Tabref}[1]{Table~\ref{tab:#1}}
\newcommand{\secref}[1]{section~\ref{sec:#1}}
\newcommand{\Secref}[1]{Section~\ref{sec:#1}}
%\newcommand{\secref}[1]{\S\ref{sec:#1}}
\newcommand{\eqnref}[1]{(\ref{eqn:#1})}
%\newcommand{\eqnref}[1]{equation~(\ref{eqn:#1})}
%\newcommand{\Eqnref}[1]{Equation~(\ref{eqn:#1})}

@@

\todo{
TODO:
\begin{itemize}
	\item Unify vocabulary and avoid synonyms.
		Sampling, grid, discretization, pixelization $\Rightarrow$ sampling (sampling schemes when not a particular set of vertices?)
		Signal, maps, fields, functions.
		Irregular, uneven, non-uniform.
		Coarsening, down-sampling.
		Sample, vertex, node, pixel, point $\Rightarrow$ vertex
	\item check English vs American English
	\item put all the code (learning experiments from Frédérick and equivariance experiments from Martino) in a common reproducible git repository to be linked from the paper
	\item update pygsp and merge sphere branch
	\item Compute the invariance error of the whole NNs (before and after training, invariance can be learned). Add it with equivariance error of the convolution to \tabref{cosmo}. \nati{I guess we are not doing that}
	\item insert \citet{jiang2019sphericalcnn} in intro: Need a global coordinate system (ok for planets, not projections like cosmo)
	\item mention that \citep{esteves2018sphericalcnn} and \citep{cohen2018sphericalcnn} computed equivariance error
\end{itemize}
}

\todo{
rebuttal:
\begin{itemize}
	\item \mdeff{Why two definitions of rotations, $R(g)$ and $\mathcal{R}_g$?} [Martino]
    \item answer commenter
	\item add somewhere (intro?): exploit symmetries to reduce sample cost $\Rightarrow$ we don't discuss symmetries much
	\item \citet{mudigonda2017climateevents} use Integrated Water Vapor (IWV) only. That most probably is the same as our TMQ, which is most probably the most informative feature. TODO: run ours with TMQ alone.
	\item Show the three samplings in a figure?
	\item better abstract (Friday's talk abstract?)
	\item conclusion: less speculation?
	\item check feedback from ICLR'20 LTS2 reading group
	\item \mdeff{check my Telegram notes}
\end{itemize}
}

@@

\todo{
Not done:
\begin{itemize}
	\item Do the experiments on the icosahedral sampling with the pooling as defined in the paper (i.e., not Jiang's specialty of taking the value of a pre-chosen sub-pixel). \nati{I guess we are not doing that.} \mdeff{I'm not sure what is the state of this, and Frédérick is not with us anymore. It's quite minor.}
	\item Check results for ModelNet40 and integrate in text. \nati{I would drop this.} \mdeff{Yeah, those results are not mature enough.}
	\item Add \citet{bruna2013gnn} in equivariance? They were the first to do a spherical CNN with graphs. Graph is built from a uniform random sampling of the sphere. \nati{Where do you want to add them?} \mdeff{They don't write how their Laplacian was built. Also, they only did MNIST on the sphere...}
\end{itemize}
}

@@

\begin{abstract}
Designing a convolution for a spherical neural network requires a delicate tradeoff between efficiency and rotation equivariance.
% exactitude with respect to equivariance.
DeepSphere, a method based on a graph representation of the sampled sphere, strikes a controllable balance between these two desiderata.
% Say something here about isotropy?
This contribution is twofold.
First, we study both theoretically and empirically how equivariance is affected by the underlying graph with respect to the number of vertices and neighbors.
Second, we evaluate DeepSphere on relevant problems.
Experiments show state-of-the-art performance and demonstrates the efficiency and flexibility of this formulation.
Perhaps surprisingly, comparison with previous work suggests that anisotropic filters might be an unnecessary price to pay.
%Say something about speed here... Not sure what?
% memory, computations, implementation complexity
\end{abstract}

@@

\begin{figure}[h]
	\centering
	\subfloat[]{
		\includegraphics[height=0.20\linewidth,valign=t]{example_brain_meg}
	} \hspace{0.1em}
	\subfloat[]{
		\includegraphics[height=0.20\linewidth,valign=t]{example_cosmo_cmb}
	} \hfill
	\subfloat[]{
		\includegraphics[height=0.20\linewidth,valign=t]{example_climate_TMQ}
		\label{fig:examples:climate}
	} \hfill
	\subfloat[]{
		\includegraphics[height=0.20\linewidth,valign=t]{example_ghcn_daily_tmax}
		\label{fig:examples:ghcn:tmax}
	} \hfill
	\subfloat[]{
		\includegraphics[height=0.20\linewidth,valign=t]{example_ghcn_graph}
		\label{fig:examples:ghcn:graph}
	}
	\caption{
		Examples of spherical data:
		(a) brain activity recorded through magnetoencephalography (MEG),\protect\footnotemark
		(b) the cosmic microwave background (CMB) temperature from \citet{planck2015overview},
		(c) hourly precipitation from a climate simulation \citep{jiang2019sphericalcnn}, %\citep{mudigonda2017climateevents},
		(d) daily maximum temperature from the Global Historical Climatology Network (GHCN).\protect\footnotemark
		A rigid full-sphere sampling is not ideal: brain activity is only measured on the scalp, the Milky Way's galactic plane masks observations, climate scientists desire a variable resolution, and the position of weather stations is arbitrary and changes over time.
		(e) Graphs can faithfully and efficiently represent sampled spherical data by placing vertices where it matters.
		%\todo{horizontal colorbar for brain}
	}
	\label{fig:examples}
\end{figure}
\footnotetext[1]{\scriptsize\url{https://martinos.org/mne/stable/auto_tutorials/plot_visualize_evoked.html}}
\footnotetext[2]{\scriptsize\url{https://www.ncdc.noaa.gov/ghcn-daily-description}}

@@

% keep the story simple and focus on the tradeoff between cost (time / computational and storage / memory) and exactitude
As neural networks (NNs) have proved to be great tools for inference, variants have been developed to handle spherical data.
Exploiting the locally Euclidean property of the sphere, early attempts used standard 2D convolutions on a grid sampling of the sphere \citep{boomsma2017sphericalcnn, su2017sphericalcnn, coors2018sphericalcnn}.
While simple and efficient, those convolutions are not equivariant to rotations.
On the other side of this tradeoff, \citet{cohen2018sphericalcnn} and \citet{esteves2018sphericalcnn} proposed to perform proper spherical convolutions through the spherical harmonic transform.
While equivariant to rotations, those convolutions are expensive (\secref{method:convolution}).

@@

As a lack of equivariance can penalize performance (\secref{exp:cosmo}) and expensive convolutions prohibit their application to some real-world problems, methods standing between these two extremes are desired.
%Such high resolutions are not necessary to represent the depth maps of 3D objects, but are required to faithfully represent cosmological and climate data.
\citet{cohen2019gauge} proposed to reduce costs by limiting the size of the representation of the symmetry group by projecting the data from the sphere to the icosahedron.
% not sure if true
The distortions introduced by this projection might however hinder performance (\secref{exp:climate}).
% (from the continuous rotation group SO(3) to a discrete rotation group)
% the dimensionality of feature maps is equal to the size of the irreducible representation of the symmetry group
%As a platonic solid, it \todo{exhibits discrete rotations}.
% the representation of its symmetry group is much smaller than SO(3)
%\todo{not all rotations, subset of SO(3)?}

Another approach is to represent the sampled sphere as a graph connecting pixels according to the distance between them \citep{bruna2013gnn, khasanova2017sphericalcnn, perraudin2019deepspherecosmo}.
While Laplacian-based graph convolutions are more efficient than spherical convolutions, they are not exactly equivariant \citep{defferrard2019deepsphereequiv}.
In this work, we argue that graph-based spherical CNNs strike an interesting balance, with a controllable tradeoff between cost and equivariance (which is linked to performance).
Experiments on multiple problems of practical interest show the competitiveness and flexibility of this approach.
% Furthermore, experiments show that, perhaps surprisingly and contrarily to previously published results \citep{cohen2019gauge} \todo{[Bronstein?]}, anisotropic filters might be an unnecessary price to pay.
% This paper focuses on improving the rotation equivariance of \citet{perraudin2019deepspherecosmo}, a scalable graph-based method designed for cosmological applications.

@@

DeepSphere leverages graph convolutions to achieve the following properties: (i) computational efficiency, (ii) sampling flexibility, and (iii) rotation equivariance (\secref{equivariance}).
% An example architecture is shown in \figref{architecture}.
The main idea is to model the sampled sphere as a graph of connected pixels: the length of the shortest path between two pixels is an approximation of the geodesic distance between them.
We use the graph CNN formulation introduced in \citep{defferrard2016graphnn} and a pooling strategy that exploits hierarchical samplings of the sphere.

\paragraph{Sampling.}
Given a sampling $\V = \{x_i \in \S^2\}_{i=1}^n$, a function $f : \S^2 \supset F_\V \to \R$ is sampled as $T_\V(f) = \b{f}$ by the sampling operator $T_\V: L^2(\S^2) \supset F_\V \to \R^n$ defined as $\b{f}: f_i=f(x_i)$, where $F_\V$ will be specified in \secref{equivariance}.
As there is no analogue of uniform sampling on the sphere, many samplings have been proposed with different tradeoffs.
In this work, depending on the considered application, we will use the equiangular \citep{driscoll1994Fouriersphere}, HEALPix \citep{gorski2005healpix}, and icosahedral \citep{baumgardner1985icosahedral} samplings.

\paragraph{Graph.}
From $\V$, we construct a weighted undirected graph $\G = (\V, w)$, where the elements of $\V$ are the vertices and the weight $w_{ij} = w_{ji}$ is a similarity measure between vertices $x_i$ and $x_j$.
The combinatorial graph Laplacian $\b{L} \in \R^{n \times n}$ is defined as $\b{L} = \b{D} - \b{A}$, where $\b{A} = (w_{ij})$ is the weighted adjacency matrix, $\b{D} = (d_{ii})$ is the diagonal degree matrix, and $d_{ii} = \sum_j w_{ij}$ is the weighted degree of vertex $x_i$.
%$\b{D} = \diag(\b{A} \b{1})$
Given a sampling $\V$, usually fixed by the application or the available measurements, the freedom in constructing $\G$ is in setting $w$.
\Secref{equivariance} shows how to set $w$ to minimize the equivariance error.
%building a good graph amounts to set a good $w$.

\paragraph{Convolution.} \label{sec:method:convolution}
On Euclidean domains, convolutions are efficiently implemented by sliding a window in the signal domain.
On the sphere however, there is no straightforward way to implement a convolution in the signal domain due to non-uniform samplings.
Convolutions are most often performed in the spectral domain through a spherical harmonic transform (SHT).
That is the approach taken by \citet{cohen2018sphericalcnn} and \citet{esteves2018sphericalcnn}, which has a computational cost of $\bO(n^{3/2})$ on isolatitude samplings (such as the HEALPix and equiangular samplings) and $\bO(n^2)$ in general.
On the other hand, following \citet{defferrard2016graphnn}, graph convolutions can be defined as
\begin{equation} \label{eqn:graph_conv}
	h(\b{L}) \b{f} = \left(\sum_{i=0}^K \alpha_i \b{L}^i\right) \b{f},
\end{equation}
where $K$ is the polynomial order (which corresponds to the filter's size) and $\alpha_i$ are the coefficients to be optimized during training.\footnote{In practice, training with Chebyshev polynomials (instead of monomials) is slightly more stable. We believe it to be due to their orthogonality and uniformity.}
Those convolutions are used by \citet{khasanova2017sphericalcnn} and \citet{perraudin2019deepspherecosmo} and cost $\bO(n)$ operations through a recursive application of $\b{L}$.\footnote{As long as the graph is sparsified such that the number of edges, i.e., the number of non-zeros in $\b{A}$, is proportional to the number of vertices $n$. This can always be done as most weights are very small.}

\paragraph{Pooling.}
Down- and up-sampling is natural for hierarchical samplings,\footnote{The equiangular, HEALPix, and icosahedral samplings are of this kind.} where each subdivision divides a pixel in (an equal number of) child sub-pixels.
To pool (down-sample), the data supported on the sub-pixels is summarized by a permutation invariant function such as the maximum or the average.
To unpool (up-sample), the data supported on a pixel is copied to all its sub-pixels.
% Frédérick: Dans le cas de l'icosahedron, le pooling n'est pas invariant à la permutation (on prend la valeur du 1er pixel, sans regarder les autres)
% Frédérick: Et le unpooling est correct uniquement dans le cas de l'average pooling (mais on fait pas de max pooling dans ce cas). Encore une fois, l'icosahedron est inspiré de Jiang, où 1 seul sub-pixel prend la valeur du pixel, le reste est rempli avec des '1'

@@

\section{Graph convolution and equivariance} \label{sec:equivariance}
%\section{Is graph convolution equivariant to rotation? [2 pages]}

@@

We require $F_\V$ to be a suitable subspace of $L^2(\S^2)$ such that $T_\V$ is invertible, i.e., the function $f \in F_\V$ can be unambiguously reconstructed from its sampled values $\b{f}$.
The existence of such a subspace depends on the sampling $\V$, and its characterization is a common problem in signal processing \citep{driscoll1994Fouriersphere}.
For most samplings, it is not known if $F_\V$ exists and hence if $T_\V$ is invertible.
A special case is the equiangular sampling where a sampling theorem holds, and thus a closed-form of $T_\V^{-1}$ is known. % \citep{driscoll1994Fouriersphere}
For samplings where no such sampling formula is available, we leverage the discrete SHT to reconstruct $f$ from $\b{f}=T_\V f$, thus approximating $T_\V^{-1}$.
For all theoretical considerations, we assume that $F_\V$ exists and $f \in F_\V$.

% 0.2 Explain why the Laplacian is so important...
By definition, the (spherical) graph convolution is rotation equivariant if and only if it commutes with the rotation operator defined as $R(g), g\in SO(3)$: $R(g) f(x) = f\left(g^{-1} x \right)$.
In the context of this work, graph convolution is performed by recursive applications of the graph Laplacian (\eqnref{graph_conv}).
Hence, if $R(g)$ commutes with $\b{L}$, then, by recursion, it will also commute with the convolution $h(\b{L})$.
As a result, $h(\b{L})$ is rotation equivariant if and only if
% 0.3 Formulate the problem
\begin{equation*} %\label{eq:equivariance}
	\b{R}_\V(g) \b{L} \b{f} = \b{L} \b{R}_\V(g) \b{f}, \hspace{1cm} \forall f\in F_\V \text{ and } \forall g\in SO(3),
\end{equation*}
where $\b{R}_\V(g) = T_\V R(g) T_\V^{-1}$.
For an empirical evaluation of equivariance, we define the \textit{normalized equivariance error} for a signal $\b{f}$ and a  rotation $g$ as
\begin{equation} \label{eq:equivariance error}
	E_{\b{L}}(\b{f}, g) = \left(\frac{ \norm {\b{R}_\V(g) \b{L} \b{f} - \b{L} \b{R}_\V(g) \b{f}} }{\norm {\b{L} \b{f}}}\right)^2.
\end{equation}
More generally for a class of signals $f \in C \subset F_\V$, the \textit{mean equivariance error} defined as
\begin{equation} \label{eq:mean equivariance error}
	\overline E_{\b{L}, C} = \mathbb E_{\b{f}\in C, g\in SO(3)} \ E_{\b{L}}(\b{f}, g)
\end{equation}
represents the overall equivariance error.
The expected value is obtained by averaging over a finite number of random functions and random rotations.

%\subsection{Finding the optimal edge weights} \label{sec:optimal}
\subsection{Finding the optimal weighting scheme} \label{sec:optimal}

% 1) summarize the work of Khasanova
Considering the equiangular sampling and graphs where each vertex is connected to 4 neighbors (north, south, east, west), \citet{khasanova2017sphericalcnn} designed a weighting scheme to minimize \eqref{eq:mean equivariance error} for longitudinal and latitudinal rotations\footnote{Equivariance to longitudinal rotation is essentially given by the equiangular sampling.}.
% because longitudinal rotation can be discrete
Their solution gives weights inversely proportional to Euclidean distances:
\begin{equation} \label{eqn:weights:khasanova}
	w_{ij} = \frac{1}{\norm{x_i-x_j}}.
\end{equation}
%leading to a high mean equivariance error if the mean is computed over rotations in SO(3).
Note that while the resulting convolution is not equivariant to the whole of $SO(3)$, that is enough for omnidirectional imaging because, as gravity consistently orients the sphere, objects only rotate longitudinally or latitudinally.

% 2) explain our approach
To achieve equivariance to all rotations, we take inspiration from \citet{belkin2005towards}.
They prove that for a \emph{random uniform sampling}, the graph Laplacian $\b{L}$ built from weights
\begin{equation} \label{eqn:weights:belkin}
	w_{ij} = e^{- \frac{1}{4t} \norm{x_i - x_j}^2}
\end{equation}
converges to the Laplace-Beltrami operator $\Delta_{\mathbb{S}^2}$ as the number of samples grows to infinity.
This result is a good starting point as $\Delta_{\mathbb{S}^2}$ commutes with rotation, i.e., $\Delta_{\mathbb{S}^2}R(g) = R(g)\Delta_{\mathbb{S}^2}$.
% and (ii) is diagonalized by the spherical harmonics.
While the weighting scheme is full (i.e., every vertex is connected to every vertex), most weights are small due to the exponential.
We hence make an approximation to limit the cost of the convolution \eqnref{graph_conv} by only considering the $k$ nearest neighbors to each vertex.
%In practice, the tricky part is to set $t$ in a non-asymptotic regime.
Given $k$, the optimal kernel width $t$ is found by searching for the minimizer of \eqref{eq:mean equivariance error}.
\Figref{kernel_widths} shows the optimal kernel widths found for various resolutions of the HEALPix sampling.
As predicted by the theory, $t_n \propto n^\beta, \beta \in \R$.
Importantly however, the optimal $t$ also depends on the number of neighbors $k$.

Considering the HEALPix sampling, \citet{perraudin2019deepspherecosmo} connected each vertex to their 8 adjacent vertices, computed the weights with \eqnref{weights:belkin}, and heuristically set $t$ to half the average squared Euclidean distance between connected vertices.
This heuristic however seems to over-estimate $t$ (see \figref{kernel_widths}).
% We observe that using a small kernel size $t$, \cite{perraudin2019deepspherecosmo} would probably be more equivariant.

\subsection{Analysis of the proposed weighting scheme}

We analyze the proposed weighting scheme both theoretically and empirically.
